---
title: "data_exploration"
author: "Christian Stenbro"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load("tidyverse",
               "rethinking", 
               "here")
```

# 1. loading experimental data

Data note: This script is using local paths, as we are not uploading the data to git... Please change this path to whatever it is on your local system.

```{r}
data <- read_csv("/Users/christianstenbro/AU/MSc/ACM/data/assignment_3_data/Simonsen_clean.csv")
```
# 2. exploring data

```{r}
summary(data)
```
Some basic questions:

What is the sample size? 

```{r}
n = length(unique(data$ID))
cat("n =", n)
```

How many trials pr. participant?

```{r}
data %>% group_by(ID) %>% count(.)
```

Total rows:

```{r}
nrow(data)
```


153 pr. participant.

```{r}
data %>% group_by(ID)
```

## 2.1 plotting variables

```{r}
data
```


```{r}
#make plots
p1 <- 
  data %>% 
  
  ggplot(aes(x = FirstRating, y = SecondRating, color = Feedback)) +
  
  geom_jitter(width = 0.5, height = 0.5, alpha = 0.2) +
  
  scale_color_continuous(high = "red", low = "blue") +
  
  geom_abline(linetype = "dashed", alpha = .7) +
  
  labs(x = "First Rating", y = "Second Rating") +
  
  ggtitle(label="Plotting first ratings against second ratings", 
          subtitle="colored by the group rating difference (feedback)") +
  
  theme_minimal()

#call plot
p1
```

```{r}
#make plots
p1 <- 
  data %>% 
  
  filter(ID %in% c(201,203,204,205,206)) %>% 
  
  ggplot(aes(x = FirstRating, y = SecondRating, color = Feedback)) +
  
  geom_jitter(width = 0.5, height = 0.5, alpha = 0.5) +
  
  scale_color_continuous(high = "red", low = "blue") +
  
  geom_abline(linetype = "dashed", alpha = .7) +
  
  labs(x = "First Rating", y = "Second Rating") +
  
  ggtitle("Plotting first ratings against second ratings for a participant subset, \ncolored by feedback") +
  
  theme_minimal() +
  
  facet_wrap(~ID, ncol = 3)

#call plot
p1
```


Comment on this plot!

Also making a binary version of this plot:

```{r}
data <- data %>% mutate(BinaryFeedback = sign(Feedback))
```


```{r}
p1 <- data %>%
  group_by(FirstRating, SecondRating) %>%
  summarise(mean_feedback = mean(Feedback), .groups = "drop") %>%
  
  ggplot(aes(x = FirstRating, y = SecondRating, fill = mean_feedback)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  geom_abline(linetype = "dashed", alpha = .5) +
  labs(x = "first rating", y = "second rating", fill = "avg feedback") +
  theme_minimal()

p1
```
## 2.2 correlation test suite

```{r}
test1 <- cor.test(x=data$FirstRating, y=data$SecondRating)
round(test1$estimate, 2)
test1
```

```{r}
test2 <- cor.test(x=data$GroupRating, y=data$SecondRating)
round(test2$estimate, 2)
```

```{r}
# just as a sanity check (this should be low.. but actually isn't; maybe it makes sense?)
test3 <- cor.test(x=data$FirstRating, y=data$GroupRating)
round(test3$estimate, 2)
```

# 3. fitting models to data

```{r}
colnames(data)
```

## 3.1 transforming the data (-1 to all ratings)

```{r}
range(data$FirstRating)
range(data$GroupRating)
range(data$SecondRating)
```


```{r}
transformed_data <- data %>% mutate(FirstRating = FirstRating - 1,
                        GroupRating = GroupRating - 1,
                        SecondRating = SecondRating - 1)
```

## 3.2 fitting wba

```{r}
# specifying model file path
stan_file <- here("wba_model.stan")
```

```{r}
# compiling Stan model
model <- cmdstan_model(stan_file,
                       cpp_options = list(stan_threads = FALSE),
                       stanc_options = list("O1"))

fit <- model$sample(
  data = list(N = nrow(transformed_data), 
                    first_rating = transformed_data$FirstRating, 
                    group_rating = transformed_data$GroupRating,
                    second_rating = transformed_data$SecondRating,
                    prior_mu = 0,
                    prior_sigma = 0.5,
                    total = rep(7, nrow(transformed_data))),      
  seed = 123,                    
  chains = 4,                    
  parallel_chains = 4,       
  iter_warmup = 1000,            
  iter_sampling = 2000,          
  refresh = 0,                   
  max_treedepth = 10,            
  adapt_delta = 0.8              
)
```

```{r}
fit
```

```{r}
saveRDS(fit, file = here("models_fit_to_data", "wba_data_model.Rdata"),
        compress = "xz")
```

## 3.3 fitting pba

```{r}
# specifying model file path
stan_file <- here("pba_model.stan")
```

```{r}
# compiling Stan model
model <- cmdstan_model(stan_file,
                       cpp_options = list(stan_threads = FALSE),
                       stanc_options = list("O1"))

fit_pba <- model$sample(
  data = list(N = nrow(transformed_data), 
                    first_rating = transformed_data$FirstRating, 
                    group_rating = transformed_data$GroupRating,
                    second_rating = transformed_data$SecondRating,
                    prior_mu = 0,
                    prior_sigma = 0.5,
                    total = rep(7, nrow(transformed_data))),      
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 0,
  max_treedepth = 10,
  adapt_delta = 0.8
)
```

```{r}
fit_pba
```
Some NAs in the ess_tail stand out... Could be a problem.!

```{r}
saveRDS(fit_pba, 
        file = here("models_fit_to_data", "pba_data_model.Rdata"),
        compress = "xz")
```

# testing that we can load the fits

```{r}
fit_pba_loaded <- readRDS(here("models_fit_to_data", "pba_data_model.Rdata"))
```

```{r}
fit_pba_loaded
```
```{r}
# checking stan stuff versioning, for collaborative debugging
packageVersion("cmdstanr")
packageVersion("rstan")
```

```{r}
draws <- fit_pba_loaded$draws()
```

```{r}
draws
```


