---
title: "prior_posterior_update_checks"
output: html_document
---

# set-up

```{r}
pacman::p_load(here, cmdstanr, tidyverse, posterior, bayesplot, ggplot2, patchwork)
```

```{r}
#Lets try this out for the PBA

file_path = here("scenario_data", "3")


# import simulation data
data_raw <- read_csv(file.path(file_path, "scenario_3_more_samples.csv"))

# import model fit
fit <- readRDS(file.path(file_path, "wba_model_scenario_3.rds"))
```
# reasoning about prior-posterior update check

The goal is to investigate the relationship between the prior and the posterior. Concretely, we ask the following:

- Has the data been informative, meaning, *is the posterior distribution visually narrower than the prior distribution?*
- Is the prior too restrictive, meaning, *is the posterior distribution's mean different from the prior distribution's mean?*

The way we do this is by visualizing the marginal posterior distribution against the marginal prior distribution. The code is based on ch. 6 in Riccardo's book:

```{r}
# First, we extract the draws from the samples
draws <- as_draws_df(fit$draws())
```

```{r}
# Then, we define the ground truth for our estimand 
true_w_first <- 0.6
true_w_group <- 0.4
true_kappa   <- true_w_first + true_w_group 
true_rho     <- true_w_first / true_kappa    
```

```{r}
# Make a helper-function that extracts the marginal prior and posterior distributions and put them in a neat tibble
make_update_df <- function(prior_vec, posterior_vec) {
  tibble(
    value        = c(prior_vec, posterior_vec),
    distribution = factor(
      rep(c("Prior", "Posterior"), each = length(prior_vec)),
      levels = c("Prior", "Posterior")
    )
  )
}
```

```{r}
#Make a plotting function 
make_update_plot <- function(df, true_val, title, xlab) {
  ggplot(df, aes(x = value, fill = distribution)) +
    geom_density(alpha = 0.6, color = "white") +
    geom_vline(xintercept = true_val, color = "black",
               linetype = "dashed", linewidth = 1) +
    scale_fill_manual(values = c("Prior" = "#ffbaba", "Posterior" = "skyblue")) +
    annotate(
      "text", x = true_val, y = Inf,
      label = sprintf("True = %.2f", true_val),
      vjust = 2,
      hjust = if (true_val > median(df$value)) 1.1 else -0.1,
      size = 4, fontface = "bold"
    ) +
    labs(
      title    = title,
      x        = xlab,
      y        = "Marginal Density"
    ) +
    theme_classic() +
    theme(
      legend.position  = "bottom",
      legend.title     = element_blank(),
      legend.text      = element_text(size = 12)
    )
}
```

```{r}
#Create a df for each parameter
df_rho     <- make_update_df(draws$rho_prior,    draws$rho)
df_kappa   <- make_update_df(draws$kappa_prior,  draws$kappa)
df_wfirst  <- make_update_df(draws$wf_prior,     draws$w_first)
df_wgroup  <- make_update_df(draws$wg_prior,     draws$w_group)
```

```{r}
#Create a plot for each parameter:
p_rho_3 <- make_update_plot(
  df_rho, true_rho,
  title = "Scenario 3 - Prior-Posterior Update: Rho",
  xlab  = "Rho"
)

p_kappa_3 <- make_update_plot(
  df_kappa, true_kappa,
  title = "Scenario 3 - Prior-Posterior Update: Kappa",
  xlab  = "Kappa"
)

```


```{r}
p_rho_3
```
```{r}
p_kappa_3
```



```{r}
#Plot!
library(patchwork)

(p_rho_1 | p_kappa_1) / (p_rho_2 | p_kappa_2) / (p_rho_3 | p_kappa_3)+
  plot_annotation(
    title = "Prior-Posterior Update Checks: Weighted Bayesian Agent - wider kappa prior",
    theme = theme(plot.title = element_text(size = 16, face = "bold"))
  )
```


Plots are interpreted in the report. One note: The prior is cut of at the limits, but I think that is expected behaviour, exactly since we are in the constrained space. However, I am surprised by how artificial this cut-off appears.


### Now do it all on real data:
```{r}
#Update the plotting function 
make_update_plot <- function(df, title, xlab) {
  ggplot(df, aes(x = value, fill = distribution)) +
    geom_density(alpha = 0.6, color = "white") +
    scale_fill_manual(values = c("Prior" = "#ffbaba", "Posterior" = "skyblue")) +
    labs(
      title    = title,
      x        = xlab,
      y        = "Marginal Density"
    ) +
    theme_classic() +
    theme(
      legend.position  = "bottom",
      legend.title     = element_blank(),
      legend.text      = element_text(size = 12)
    )
}
```

```{r}
#WBA

draws_list <- readRDS(here("models_fit_to_data", "wba_empirical_draws.rds"))

selected_ids <- c("203", "225", "242")

for (id in selected_ids) {
  
  draws <- draws_list[[id]]
  
  # Build update dfs
  df_rho    <- make_update_df(draws$rho_prior,   draws$rho)
  df_kappa  <- make_update_df(draws$kappa_prior, draws$kappa)
  df_wfirst <- make_update_df(draws$wf_prior,    draws$w_first)
  df_wgroup <- make_update_df(draws$wg_prior,    draws$w_group)
  
  # Make plots
  p_rho    <- make_update_plot(df_rho,    title = "Rho",    xlab = "Rho")
  p_kappa  <- make_update_plot(df_kappa,  title = "Kappa",  xlab = "Kappa")
  p_wfirst <- make_update_plot(df_wfirst, title = "Wfirst", xlab = "wfirst")
  p_wgroup <- make_update_plot(df_wgroup, title = "Wgroup", xlab = "wgroup")
  
  print(
    (p_rho | p_kappa) / (p_wfirst | p_wgroup) +
      plot_annotation(
        title = paste("Prior-Posterior Update: Participant", id),
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
  )
}
```

```{r}
#PBA 
draws_list <- readRDS(here("models_fit_to_data", "pba_empirical_draws.rds"))

selected_ids <- c("203", "225", "242")

for (id in selected_ids) {
  
  draws <- draws_list[[id]]
  
  # Build update dfs
  df_wfirst <- make_update_df(draws$wf_prior,    draws$w_first)
  df_wgroup <- make_update_df(draws$wg_prior,    draws$w_group)
  
  # Make plots
  p_wfirst <- make_update_plot(df_wfirst, title = "Wfirst", xlab = "wfirst")
  p_wgroup <- make_update_plot(df_wgroup, title = "Wgroup", xlab = "wgroup")
  
  print(
    (p_wfirst | p_wgroup) +
      plot_annotation(
        title = paste("PBA - Prior-Posterior Update: Participant", id),
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
  )
}
```



```{r}
# Extract all posterior draws for wfirst and wgroup
draws_list <- readRDS(here("models_fit_to_data", "wba_empirical_draws.rds"))
all_ids <- names(draws_list)

posterior_df <- map_dfr(all_ids, function(id) {
  draws <- draws_list[[id]]
  tibble(
    id       = id,
    w_first  = draws$w_first,
    w_group  = draws$w_group
  )
})

# Color palette — one color per participant
n_ids <- length(all_ids)
colors <- rainbow(n_ids)
names(colors) <- all_ids

# Plot wfirst
p_all_wfirst <- ggplot(posterior_df, aes(x = w_first, color = id)) +
  geom_density(linewidth = 0.2, alpha = 0.8) +
  scale_color_manual(values = colors, guide = "none") +
  xlim(0, 5) +
  ylim(0, 15) +
  labs(
    x     = "w_first",
    y     = "Density",
    color = "Participant"
  ) +
  theme_minimal()
    

# Plot wgroup
p_all_wgroup <- ggplot(posterior_df, aes(x = w_group, color = id)) +
  geom_density(linewidth = 0.2, alpha = 0.8) +
  scale_color_manual(values = colors, guide = "none") +
  xlim(0, 5) +
  ylim(0, 15) +
  labs(
    x     = "w_group",
    y     = "Density",
  ) +
  theme_minimal()
    
print(
    (p_all_wfirst | p_all_wgroup) +
      plot_annotation(
        title = paste("WBA - Posterior Density: all participants"),
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      ))
```


```{r}
# Extract all posterior draws for wfirst and wgroup
draws_list <- readRDS(here("models_fit_to_data", "pba_empirical_draws.rds"))
all_ids <- names(draws_list)

posterior_df <- map_dfr(all_ids, function(id) {
  draws <- draws_list[[id]]
  tibble(
    id       = id,
    w_first  = draws$w_first,
    w_group  = draws$w_group
  )
})

# Color palette — one color per participant
n_ids <- length(all_ids)
colors <- rainbow(n_ids)
names(colors) <- all_ids

# Plot wfirst
p_all_wfirst <- ggplot(posterior_df, aes(x = w_first, color = id)) +
  geom_density(linewidth = 0.2, alpha = 0.8) +
  scale_color_manual(values = colors, guide = "none") +
  xlim(0, 1) +
  ylim(0, 10) +
  labs(
    x     = "w_first",
    y     = "Density",
    color = "Participant"
  ) +
  theme_minimal()
    

# Plot wgroup
p_all_wgroup <- ggplot(posterior_df, aes(x = w_group, color = id)) +
  geom_density(linewidth = 0.2, alpha = 0.8) +
  scale_color_manual(values = colors, guide = "none") +
  xlim(0, 1) +
  ylim(0, 10) +
  labs(
    x     = "w_group",
    y     = "Density",
  ) +
  theme_minimal()
    
print(
    (p_all_wfirst | p_all_wgroup) +
      plot_annotation(
        title = paste("PBA - Posterior Density: all participants"),
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      ))

```

