---
title: "wba"
author: "Christian Stenbro"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# write up of the task we are simulating

The goal of this script is to make a simulated agent that mimics human behaviour in the social conformity experimental task. The task is described in Simonsen et al. (2019).

The task proceeds in two rounds. In the first, participants are asked to rate the trustworthiness of the people depicted in *153* face images. The ratings are given on a scale from 1 (not trustworthy at all) to 8 (very trustworthy).

After rating each image, participants are shown a score communicating a (made up) 'group' rating of trustworthiness for the given image. Upon rating all images, participants take a 1 hour break before the second round, in which they have to rate the same images. 

The goal of the experiment is to assess to which extent participants second rate is swayed by the group rating.

Regarding the made up group ratings, the author's write the following: "In order to provide an adequate number of trials in each condition, and minimize the number of participants required in the study, others’ ratings were generated pseudo-randomly by the computer with the following cri- teria: in approximately 1/3 of the trials, these ratings agreed with the participant’s rating, in approximately 1/3 ratings were two or three points lower than the participant’s, and in approximately 1/3 ratings were two or three points higher."

*References*
Simonsen, A., Fusaroli, R., Skewes, J. C., Roepstorff, A., Mors, O., Bliksted, V., & Campbell-Meiklejohn, D. (2019). Socially Learned Attitude Change is not reduced in Medicated Patients with Schizophrenia. Scientific Reports, 9(1), 992. https://doi.org/10.1038/s41598-018-37250-x

# setting up a weighted bayesian agent

I am thinking about using the following:

```{r}
posterior = Beta(
  alpha_0 + w_1 * k_1 + w_2 * k_2,
  beta_0 + w_1 * (n-1 - k_1) + w_2 * (n_2 - k_2)
)

# exactly what was k and w? w are weights. And k is, probably the first rating (?) and the group rating (?) --> though non of these are remembered consciously, so I wonder how well this would actually fit the data? Would it make sense to include some noise here?
```

Understanding the variables:

k_1 = own rating from round 1 or *direct evidence*
k_2 = group rating or *social evidence*

I do wonder if it makes sense to call k_1 direct evidence? It is almost more like the prior belief..?

First, I will set up some placeholder input variable vectors:

```{r}
rounds <- 10

round_1_rating <- as.integer(runif(rounds, 1, 8))
group_rating <- as.integer(runif(rounds, 1, 8))
```

We want an agent to generate new ratings based on these inputs, weighing them (?) against each other.


